An IFS study of Low-powered Radio Galaxies

### 2019-12-21 paper.tex
Work on Ionsied gas section
 On branch master
 Changes to be committed:
	modified:   paper.tex

We also make use of the Atlas$^\text{3D}$ specific angular momentum
parameter, $\lambda_\text{R}$ \citep{Emsellem2007}, evaluated at
$1$~$R_\text{e}$, where
\begin{equation}
  \lambda_\mathrm{R_e}\equiv\frac{\sum_{i=1}^{N}\,F_iR_i|V_i|}{\sum_{i=1}^{N}\,F_iR_i\sqrt{V_i^2+\sigma_i^2}}~,
\end{equation}
{\bf (Is $R_i$ really measured from the centre as stated, or rather
  with respect to the kinematic minor axis/rotation axis?
  State/explain if the latter. JW: $R_i$ is from the center - see Eq6 in \citep{Emsellem2007})} where $F_i$ is the flux of the
$i^\text{th}$ bin, $R_i$ its distance from the centre, $V_i$ its mean
stellar velocity and $\sigma_i$ its stellar velocity dispersion. The
sum is computed over all $N$ bins within \removed{$1$~$R_\text{e}$} \added{$1$~$R_\text{max}$} of the
centre of the galaxy. {\bf (What do you do if the field of view does
  not reach $1$~$R_\text{e}$? Specify/discuss (as for
  $\epsilon_\text{e}$).)}

@@

We find $4$ out of $11$, or $36\pm15\%$, of our sample galaxies are
regular rotators. This compares to $82\%$ of the Atlas$^\text{3D}$
galaxies being classified as regular rotators
\citep{Krajnovic2011}. {\bf (Better to also quote the uncertainty on
  that $82\%$ figure.)} However, the likelihood of a given galaxy
being a regular rotator is highly mass dependent, the likelihood
decreasing with increasing mass. As our Southern Sample galaxies are
all extremely massive, we naturally expect to find a higher fraction
of non-regular rotators than that of the Atlas$^\text{3D}$ sample.

@@

We find $5$ out of $11$ galaxies, or $45\pm13\%$, are slow rotators
(see Fig.~\ref{fig:lambdaR_ellip}). This fraction is between that of
the Atlas$^\text{3D}$ and MASSIVE surveys, that find respectively
$13.1\%$ and $77.5\%$ of their sample galaxies to be slow
rotators. {\bf (Again, better to also quote the uncertainties on those
  two fractions. I doubt the decimal place is meaningful in either of
  them.)} As in the regular-/non-regular-rotator classification scheme
(Section~\ref{subsec:RR-NRR}), slow rotators are more likely to be
found in higher mass galaxies, and the three samples
(Atlas$^\text{3D}$, MASSIVE and our Southern Sample) have very
different mass distributions (see the upper panel of
Fig.~\ref{fig:SRmassFraction}).

@@

\citet{Bender1993} observed a tight relationship between the global
Mg$_2$ absorption line strength (a variant of the Mg~$b$ line
strength) and the logarithm of the central stellar velocity
dispersion, $\sigma_0$, in elliptical galaxies. They also noticed that
the residuals from the best-fitting linear relation appeared to be
intrinsic, with a near Gaussian distribution about the median value
{\bf (Was that median value $\approx0$? If so, state explicitly. If
  not, then the fit was either bad or there is something we do not
  understand.)} and no correlation with any other galaxy property.

Part of the red continuum band of the Mg$_2$ index is outside the
spectral range of VIMOS, but \citet{Ziegler1997} showed there is an
approximately linear relationship between Mg~$b$ and Mg$_2$, with a
gradient of $14.3$\,--\,$15.5$~mag~\AA$^{-1}$ depending on
metallicity. {\bf (Why the strange units? Is Mg$_2$ measured in mag
  rather than \AA? If so, add a short note. If not, correct.)} Thus,
following \citet{Ziegler1997}, we also use Mg~$b$ as a proxy for
Mg$_2$.

Using a central circular aperture with a radius of $2\arcsec$ in each
datacube, we measure $\sigma_0$ and Mg~$b$ at a degraded spectral
resolution of $8.4$~\AA\ full-width at half-maximum (FWHM), chosen
such that the transformation functions of \citet{Vazdekis2010} can be
used to change literature measurements from the Lick/IDS system to the
LIS. The results are shown in Fig.~\ref{fig:globalMg}, with the
best-fitting linear relationship
% (found using \textsc{lts\_linefit} from \citealt{Cappellari2013})
indicated by the solid black line, with a gradient of
$3.3\pm1.0$~\AA~dex$^{-1}$. It was noted that NGC~1316 is a possible
outlier. {\bf (Noted by who or where? Specify. Also mention that it
  seems to again be an outlier here.)} Excluding NGC~1316 from the fit
yields a gradient of $2.5\pm0.7$~\AA~dex$^{-1}$ (green dashed line in
Fig.~\ref{fig:globalMg}). Both fits are in very good agreement with
the gradient found by \citet{Ziegler1997} for a sample of elliptical
galaxies in the Virgo and Coma clusters of galaxies
% using data from \citet{Dressler1987}
($2.7$~\AA~dex$^{-1}$, transformed to the LIS system using the
functions of \citealt{Vazdekis2010}; blue dotted line in
Fig.~\ref{fig:globalMg}). {\bf (Any uncertainty on this value of
  $2.7$? If so, quote it.)}
% , with an intrinsic scatter of 0.16\,\AA\ for galaxies with
% $\log \sigma_0 \geqslant 2.3$\,dex.

{\bf (One element I find confusing in this sub-section is that the
  first paragraph mentions a ``global'' relation, but the third
  paragraph and Fig.~3 use central $2\arcsec$ radius apertures that
  are therefore ``central'' apertures. Which one is right? You need to
  harmonise this. Also, in Fig.~\ref{fig:globalMg}, are you using the
  full relation of \citet{Ziegler1997} or just their gradient (and you
  then select a zero-point that suits you)? If the latter, then you
  must state this in the text. In other words, is the near perfect
  match of the green dashed and blue dotted lines truly a result or is
  it partially by construction (i.e.\ because you matched the
  zero-points yourself)? Clarify.)}

@@

Table \ref{tab:litAbsorption} summarises our comparisons to multiple
sets of literature measurements, where the offsets and dispersions
listed are defined as the means and standard deviations of the
differences between the literature measurements and ours. For the
comparison to \citet{Vazdekis2010}, we re-analysed the Spectral Aeral
Unit for Research on Optical Nebulae (SAURON)
dataset\footnote{\url{http://www.strw.leidenuniv.nl/sauron/}}
\citep{Emsellem2004} using the same analysis pipeline as for our
Southern Sample. {\bf (Why not use the much larger Atlas$^{\rm 3D}$
  dataset instead? Explain briefly if justified.)}

@@

\citet{Kuntschner2010} showed that KDCs can be separated into two
classes: either small with various stellar ages or typically larger
but containing a homogeneously old stellar population. {\bf (I thought
  the first to point this out was McDermid et al.\ (2006; see e.g.\
  their Fig.~16), which should at least also be quoted here. Check and
  possibly correct.)} The size of the KDC is defined as the radius
where the superposition of the two kinematic components (the KDC and
the host galaxy) results in a local minimum in the mean velocity. The
age of the most-likely SSP model for the spatially integrated spectrum
within a central aperture of $1\arcsec$ {\bf (Again, square or
  circular, radius or diameter? Clarify.)} is taken as the KDC's
age. {\bf (Are these two definitions (of the KDCs' sizes and ages)
  those of \citealt{Kuntschner2010} or yours or both? Clarify.)}

@@

Only $4$ galaxies (IC~1459, NGC~612, NGC~1316 and NGC~3100) of our
Southern Sample have emission lines detected outside of their central
region. Of the remaining $7$ galaxies, $5$ (IC~1531, IC~4296,
NGC~3557, NGC~7075 and ESO~443-G024) are only detected in their centre
(so only the integrated flux map is shown in Appendix~\ref{sec:maps}),
and $2$ (NGC~1399 and PKS~718-34) have no detection of any emission
line in a spatially-resolved manner (so no ionised gas map is shown in
Appendix~\ref{sec:maps}). For each individual galaxy, all the emission
lines have a similar distribution (but different fluxes). {\bf (It
  would help in this paragraph to define what you mean by
  ``centre''. E.g.\ ``a centred circular region of radius
  $???\arcsec$'' or ``within $???\arcsec$ of the centre. Add.)}

As can be seen from the maps in Appendix~\ref{sec:maps}, both NGC~612
and NGC~3100 have significant off-centred ionised gas emission. {\bf
  (What do you mean by ``significant'' here? Bright?
  Clarify. Equivalently, why is the off-centred emission in IC~1459
  and NGC~1316 not considered significant?)} In NGC~612, the ionised
gas does not appear related to either the radio jet (extending
$\approx200$~kpc to the East and West of the galaxy centre) or the CO,
but in NGC~3100 the ionised gas is spatially coincident with both the
radio jet and the gap in the CO ring.

@@

Our calculated ionised gas masses are listed in
Table~\ref{tab:derivedProp}. They are at the upper edge of, and
possibly exceed, {\bf (Why ``possibly''? They either always exceed,
  generally exceed, occasionally exceed, or never exceed! Check and
  clarify.)} the typical masses observed in ETGs
\citep[e.g.][]{Phillips1986, Zeilinger1996, Sarzi2005}. Given that our
sample contains particularly massive galaxies, however, high ionised
gas masses are not unexpected.

@@

\textbf{IC~1459} has ionised gas counter-rotating with respect to the
stellar KDC, but it is also misaligned with respect to the outer stars
of the host galaxy (i.e.\ the stars beyond the KDC), that are
non-rotating (see Fig.~\ref{fig:ic1459}). The ionised gas thus appears
to have a third, independent, axis of rotation. {\bf (I do not
  understand why you state/claim that here. From the map shown, it
  seems pretty clear to me that the stars in the outer parts of the
  galaxy {\em are} rotating (i.e.\ the top-left is red and the
  bottom-right is blue), and that they are rotating in a manner
  consistent with the ionised gas, just as \citet{Franx1988}
  claimed. Convince me otherwise of change your statement!)} This is
contrary to \citet{Franx1988} who claims that the rotation of the gas
is aligned with that of the stars outside of the KDC. As the stellar
disc did not survive the merger event that created the KDC, {\bf (How
  do you know there was a stellar disc to start with? Clarify or
  rephrase.)} it is unlikely that the gas in the progenitor of IC~1459
survived. However, the lost gas may have remained bound and been
re-accreted, so that the gas observed today preserves a fossil record
of the galaxy's original disc. Alternatively, the ISM may have yet
another external origin.

@@

In terms of absorption line strengths and stellar populations
(Sections~\ref{sec:absorption} and \ref{sec:stellarPop}), we found for
our Southern Sample galaxies a gradient of the central Mg~$b$ line
strength versus stellar velocity dispersion relation remarkably
similar to that of \citet{Ziegler1997}. Most of the galaxies in our
Southern Sample have typical ETG properties, with old, metal-rich
stellar populations. {\bf (What about $\alpha$-elements? You should
  also summarise those results here.)} The few exceptions are NGC~612
and NGC~1316, that are both significantly younger than most ETGs, and
NGC~3557, that has a young core. All $3$ KDCs contain old stellar
populations, and they are consistent with being merger remnants. The
KDCs found in Section~\ref{sec:StarKine} are also in agreement with
the KDC age\,--\,size relation found by \citet{Kuntschner2010} in
optically-selected ETGs (see Section~\ref{subsec:popKDC}). {\bf (See
  my previous comment about McDermid et al.\ 2006, and perhaps quote
  it here instead or additionally.)}

>>>>

We also make use of the Atlas$^\text{3D}$ specific angular momentum
parameter, $\lambda_\text{R}$ \citep{Emsellem2007}, evaluated at
$1$~$R_\text{e}$, where
\begin{equation}
  \lambda_\mathrm{R_e}\equiv\frac{\sum_{i=1}^{N}\,F_iR_i|V_i|}{\sum_{i=1}^{N}\,F_iR_i\sqrt{V_i^2+\sigma_i^2}}~,
\end{equation}
{\bf (Is $R_i$ really measured from the centre as stated, or rather
  with respect to the kinematic minor axis/rotation axis?
  State/explain if the latter. JW: Yes, $R_i$ is from the center - see Eq6 in \citep{Emsellem2007})} where $F_i$ is the flux of the
$i^\text{th}$ bin, $R_i$ its distance from the centre, $V_i$ its mean
stellar velocity and $\sigma_i$ its stellar velocity dispersion. The
sum is computed over all $N$ bins within \removed{$1$~$R_\text{e}$} \added{$1$~$R_\text{max}$} of the
centre of the galaxy. {\bf (What do you do if the field of view does
  not reach $1$~$R_\text{e}$? Specify/discuss (as for
  $\epsilon_\text{e}$).)}

@@

We find $4$ out of $11$, or $36\pm15\%$, of our sample galaxies are
regular rotators. This compares to \added{$82\pm2\%$} of the Atlas$^\text{3D}$
galaxies being classified as regular rotators
\citep{Krajnovic2011}. {\bf (Better to also quote the uncertainty on
  that $82\%$ figure.)} However, the likelihood of a given galaxy
being a regular rotator is highly mass dependent, the likelihood
decreasing with increasing mass. As our Southern Sample galaxies are
all extremely massive, we naturally expect to find a higher fraction
of non-regular rotators than that of the Atlas$^\text{3D}$ sample.

@@

We find $5$ out of $11$ galaxies, or $45\pm13\%$, are slow rotators
(see Fig.~\ref{fig:lambdaR_ellip}). This fraction is between that of
the Atlas$^\text{3D}$ and MASSIVE surveys, that find respectively
\added{$13\pm2\%$} and \added{$80\pm6\%$} of their sample galaxies to be slow
rotators. {\bf (Again, better to also quote the uncertainties on those
  two fractions. I doubt the decimal place is meaningful in either of
  them.)} As in the regular-/non-regular-rotator classification scheme
(Section~\ref{subsec:RR-NRR}), slow rotators are more likely to be
found in higher mass galaxies, and the three samples
(Atlas$^\text{3D}$, MASSIVE and our Southern Sample) have very
different mass distributions (see the upper panel of
Fig.~\ref{fig:SRmassFraction}).

@@

\citet{Bender1993} observed a tight relationship between the \added{central} \removed{global}
Mg$_2$ absorption line strength (a variant of the Mg~$b$ line
strength) and the logarithm of the central stellar velocity
dispersion, $\sigma_0$, in elliptical galaxies. \removed{They also noticed that
the residuals from the best-fitting linear relation appeared to be
intrinsic, with a near Gaussian distribution about the median value
\textbf{(Was that median value $\approx0$? If so, state explicitly. If
  not, then the fit was either bad or there is something we do not
  understand.)} and no correlation with any other galaxy property.}

Part of the red continuum band of the Mg$_2$ index is outside the
spectral range of VIMOS, but \citet{Ziegler1997} showed there is an
approximately linear relationship between Mg~$b$ and Mg$_2$, with a
gradient of $14.3$\,--\,$15.5$~mag~\AA$^{-1}$ depending on
metallicity. {\bf (Why the strange units? Is Mg$_2$ measured in mag 
rather than \AA? If so, add a short note. If not, correct.)} 
\added{(Mg$_2$ is condered to be a molecular index and is thus measured 
in units of mag$^{-1}$.) We can thus}\removed{Thus, following \citet{Ziegler1997}, we also} use Mg~$b$ as a proxy for
Mg$_2$.

Using a central circular aperture with a radius of $2\arcsec$ in each
datacube, we measure $\sigma_0$ and Mg~$b$ at a degraded spectral
resolution of $8.4$~\AA\ full-width at half-maximum (FWHM), chosen
such that the transformation functions of \citet{Vazdekis2010} can be
used to change literature measurements from the Lick/IDS system to the
LIS. The results are shown in Fig.~\ref{fig:globalMg}, with the
best-fitting linear relationship
% (found using \textsc{lts\_linefit} from \citealt{Cappellari2013})
indicated by the solid black line, with a gradient of
$3.3\pm1.0$~\AA~dex$^{-1}$. \added{We} \removed{It was} noted that NGC~1316 is \added{again an} \removed{a possible}
outlier. {\bf (Noted by who or where? Specify. Also mention that it
  seems to again be an outlier here.)} Excluding NGC~1316 from the fit
yields a gradient of $2.5\pm0.7$~\AA~dex$^{-1}$ (green dashed line in
Fig.~\ref{fig:globalMg}). Both fits are in very good agreement with
the gradient found by \citet{Ziegler1997} for a sample of elliptical
galaxies in the Virgo and Coma clusters of galaxies
% using data from \citet{Dressler1987}
($2.7$~\AA~dex$^{-1}$, transformed to the LIS system using the
functions of \citealt{Vazdekis2010}; blue dotted line in
Fig.~\ref{fig:globalMg}). {\bf (Any uncertainty on this value of
  $2.7$? If so, quote it.) JW: No \citet{Ziegler1997} doesn't propagate uncertainty into errors on the fit.}
% , with an intrinsic scatter of 0.16\,\AA\ for galaxies with
% $\log \sigma_0 \geqslant 2.3$\,dex.

{\bf (One element I find confusing in this sub-section is that the
  first paragraph mentions a ``global'' relation, but the third
  paragraph and Fig.~3 use central $2\arcsec$ radius apertures that
  are therefore ``central'' apertures. Which one is right? You need to
  harmonise this. Also, in Fig.~\ref{fig:globalMg}, are you using the
  full relation of \citet{Ziegler1997} or just their gradient (and you
  then select a zero-point that suits you)? If the latter, then you
  must state this in the text. In other words, is the near perfect
  match of the green dashed and blue dotted lines truly a result or is
  it partially by construction (i.e.\ because you matched the
  zero-points yourself)? Clarify.) JW: the entire function has been transformed including the zero-point. This actually results in a cubic function (see Appendix A of \citealt{Vazdekis2010}), though as can been seen from the Fig.~3, the quadradic and cubic terms have very little effect in this range.}

@@

Table \ref{tab:litAbsorption} summarises our comparisons to multiple
sets of literature measurements, where the offsets and dispersions
listed are defined as the means and standard deviations of the
differences between the literature measurements and ours. For the
comparison to \citet{Vazdekis2010}, we re-analysed the Spectral Aeral
Unit for Research on Optical Nebulae (SAURON)
dataset\footnote{\url{http://www.strw.leidenuniv.nl/sauron/}}
\citep{Emsellem2004} using the same analysis pipeline as for our
Southern Sample. {\bf (Why not use the much larger Atlas$^{\rm 3D}$
  dataset instead? Explain briefly if justified.)} \added{This was used over the larger Atlas$^\text{3D}$ sample \citep{Cappellari2011}, as stellar populations were not published by the Atlas$^\text{3D}$ team.}

@@

\added{\citet{McDermid2006}} \removed{\citet{Kuntschner2010}} showed that 
KDCs can be separated into two classes: either small with various stellar 
ages or typically larger but containing a homogeneously old stellar 
population. {\bf (I thought the first to point this out was 
McDermid et al.\ (2006; see e.g.\ their Fig.~16), which should at 
least also be quoted here. Check and possibly correct.)} 
\added{Using the method of \ctiet{McDermid2006},} the size 
\added{and age} of the KDC is defined as the radius
where the superposition of the two kinematic components (the KDC and
the host galaxy) results in a local minimum in the mean velocity \added{and} the
age of the most-likely SSP model for the spatially integrated spectrum
within a central \added{, circular} aperture with a \added{radius} of 
$1\arcsec$ {\bf (Again, square or circular, radius or diameter? Clarify.)} 
\removed{is taken as the KDC's age.} \added{, repectively.} 
{\bf (Are these two definitions (of the KDCs' sizes and ages) 
those of \citealt{Kuntschner2010} or yours or both? Clarify.)}

@@

Only $4$ galaxies (IC~1459, NGC~612, NGC~1316 and NGC~3100) of our
Southern Sample have emission lines detected outside of a central
\removed{region} \added{1--2 spaxels. This means no structure can be
 resolved}. Of the remaining $7$ galaxies, $5$ (IC~1531, IC~4296,
NGC~3557, NGC~7075 and ESO~443-G024) are only detected in their centre
(so only the integrated flux map is shown in Appendix~\ref{sec:maps}),
and $2$ (NGC~1399 and PKS~718-34) have no detection of any emission
line in a spatially-resolved manner (so no ionised gas map is shown in
Appendix~\ref{sec:maps}). For each individual galaxy, all the emission
lines have a similar distribution (but different fluxes). {\bf (It
  would help in this paragraph to define what you mean by
  ``centre''. E.g.\ ``a centred circular region of radius
  $???\arcsec$'' or ``within $???\arcsec$ of the centre. Add.)}

\removed{As can be seen from the maps in Appendix~\ref{sec:maps}, 
both NGC~612 and NGC~3100 have significant off-centred ionised 
gas emission.} {\bf (What do you mean by ``significant'' here? Bright?
  Clarify. Equivalently, why is the off-centred emission in IC~1459
  and NGC~1316 not considered significant?)} 
\added{The ionised gas emission for both NGC~612 ).} In NGC~612, 
the ionised gas does not appear related to either the radio jet 
(extending $\approx200$~kpc to the East and West of the galaxy 
centre) or the CO, but in NGC~3100 the ionised gas is spatially 
coincident with both the radio jet and the gap in the CO ring.

@@

Our calculated ionised gas masses are listed in
Table~\ref{tab:derivedProp}. They are at the upper \removed{edge} 
\added{range} of, and \removed{possibly} \added{occasionally} exceed, 
{\bf (Why ``possibly''? They either always exceed, generally exceed, 
occasionally exceed, or never exceed! Check and clarify.)} the 
typical masses observed in ETGs 
\citep[e.g.][]{Phillips1986, Zeilinger1996, Sarzi2005}. Given that 
our sample contains particularly massive galaxies, however, 
high ionised gas masses are not unexpected.

@@

\removed{\textbf{IC~1459} has ionised gas counter-rotating with respect to the
stellar KDC, but it is also misaligned with respect to the outer stars
of the host galaxy (i.e.\ the stars beyond the KDC), that are
non-rotating (see Fig.~\ref{fig:ic1459}). The ionised gas thus appears
to have a third, independent, axis of rotation. {\bf (I do not
  understand why you state/claim that here. From the map shown, it
  seems pretty clear to me that the stars in the outer parts of the
  galaxy {\em are} rotating (i.e.\ the top-left is red and the
  bottom-right is blue), and that they are rotating in a manner
  consistent with the ionised gas, just as \citet{Franx1988}
  claimed. Convince me otherwise of change your statement!)} This is
contrary to \citet{Franx1988} who claims that the rotation of the gas
is aligned with that of the stars outside of the KDC. As the stellar
disc did not survive the merger event that created the KDC, {\bf (How
  do you know there was a stellar disc to start with? Clarify or
  rephrase.)} it is unlikely that the gas in the progenitor of IC~1459
survived. However, the lost gas may have remained bound and been
re-accreted, so that the gas observed today preserves a fossil record
of the galaxy's original disc. Alternatively, the ISM may have yet
another external origin.}

\added{\textbf{IC~1459} has ionised gas co-rotating with the outer part (not the KDC) of the host galaxy \citep{Franx1988}. Whilst this is consistent with an internal origin of the gas, it should be noted that a major merger is typically expected to deplete a galaxy's gas reservoir. Either this did not happen in this case (a major merger being evidenced by the existence of the KDC) or the ISM may has an external origin and the apparent alignment of the gas and outer stars is random.}

@@

In terms of absorption line strengths and stellar populations
(Sections~\ref{sec:absorption} and \ref{sec:stellarPop}), we found for
our Southern Sample galaxies a gradient of the central Mg~$b$ line
strength versus stellar velocity dispersion relation remarkably
similar to that of \citet{Ziegler1997}. Most of the galaxies in our
Southern Sample have typical ETG properties, with old, metal-rich
stellar populations. {\bf (What about $\alpha$-elements? You should
  also summarise those results here.)} The few exceptions are NGC~612
and NGC~1316, that are both significantly younger than most ETGs, and
NGC~3557, that has a young core. All $3$ KDCs contain old stellar
populations, and they are consistent with being merger remnants. The
KDCs found in Section~\ref{sec:StarKine} are also in agreement with
the KDC age\,--\,size relation found by \added{\citet{McDermid2006}} 
\removed{\citet{Kuntschner2010}} in optically-selected ETGs (see 
Section~\ref{subsec:popKDC}). {\bf (See my previous comment about 
McDermid et al.\ 2006, and perhaps quote it here instead or 
additionally.)}
